\documentclass[11pt,a4paper,fleqn]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[margin=2.5cm]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{array,longtable}
\usepackage{graphicx}
\usepackage{float}
\usepackage{mathptmx}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,calc}
\usepackage[hidelinks]{hyperref}

\setlength{\mathindent}{0pt}

\title{Spot-Market BESS EPEC Extension}
\author{Author}
\date{\today}

\begin{document}

\maketitle

\section*{Notation}

\noindent The revised extension focuses only on deterministic day-ahead spot-market operation.

\begingroup
\renewcommand{\arraystretch}{1.15}
\setlength{\LTleft}{0pt}
\setlength{\LTright}{0pt}
\begin{longtable}{@{}p{0.22\textwidth}
                  @{\hspace{0.015\textwidth}}p{0.55\textwidth}
                  @{\hspace{0.015\textwidth}}p{0.18\textwidth}@{}}
\hline
\textbf{Symbol} & \textbf{Description} & \textbf{Unit} \\
\hline
\endfirsthead
\hline
\textbf{Symbol} & \textbf{Description} & \textbf{Unit} \\
\hline
\endhead
\hline
\endfoot
\multicolumn{3}{@{}l}{\textbf{Sets and Indices}} \\
\hline
\(t \in \mathcal{T}\) & Time-period index and set. & -- \\
\(\tau \in \mathcal{T}^{SOC}\) & State-of-charge time index, including the initial state. & -- \\
\(g \in \mathcal{G}\) & Generator index and set. & -- \\
\(n \in \mathcal{N}\) & Network node index and set. & -- \\
\(G_n\) & Set of generators located at node \(n\). & -- \\
\(i \in \mathcal{I}\) & BESS investor or storage-unit index and set. & -- \\
\(l \in \mathcal{L}\) & Transmission-line index and set. & -- \\
\hline
\multicolumn{3}{@{}l}{\textbf{Upper-Level Decision Variables}} \\
\hline
\(X^{power}_{i,n}\) & Installed BESS power capacity chosen by investor \(i\) at node \(n\). & MW \\
\(X^{energy}_{i,n}\) & Installed BESS energy capacity chosen by investor \(i\) at node \(n\). & MWh \\
\hline
\multicolumn{3}{@{}l}{\textbf{Lower-Level Decision Variables}} \\
\hline
\(P_{g,t}\) & Dispatch of generator \(g\) in time period \(t\). & MW \\
\(P^{shed}_{n,t}\) & Load shedding at node \(n\) and time \(t\). & MW \\
\(NetInjection_{n,t}\) & Net nodal injection used in the DC network representation. & MW \\
\(P^{charge}_{i,n,t}\) & BESS charging power. & MW \\
\(P^{discharge}_{i,n,t}\) & BESS discharging power. & MW \\
\(SOC_{i,n,\tau}\) & State of charge of BESS \(i\) at node \(n\) and SOC time \(\tau\). & MWh \\
\hline
\multicolumn{3}{@{}l}{\textbf{Dual Variables}} \\
\hline
\(\lambda_{n,t}\) & Dual variable of the nodal net-injection definition. & \texteuro{}/MWh \\
\(\lambda_{sys,t}\) & Dual variable of system-wide power balance. & \texteuro{}/MWh \\
\(\nu^-_{g,t},\nu^+_{g,t}\) & Dual variables of the lower and upper generator dispatch bounds. & \texteuro{}/MWh \\
\(\mu^-_{l,t},\mu^+_{l,t}\) & Dual variables of the lower and upper transmission-flow bounds. & \texteuro{}/MWh \\
\(\rho^+_{i,n,t}\) & Dual variable of the BESS charging power-capacity bound. & \texteuro{}/MWh \\
\(\sigma^+_{i,n,t}\) & Dual variable of the BESS discharging power-capacity bound. & \texteuro{}/MWh \\
\(\gamma_{i,n,t}\) & Dual variable of the SOC transition equation. & \texteuro{}/MWh \\
\(\delta^-_{i,n,\tau},\delta^+_{i,n,\tau}\) & Dual variables of the lower and upper SOC energy bounds. & \texteuro{}/MWh \\
\(\xi^-_{n,t},\xi^+_{n,t}\) & Dual variables of the lower and upper load-shedding bounds. & \texteuro{}/MWh \\
\(\lambda^{cap}_n\) & Nodal BESS connection-capacity price (planned extension): dual of the shared access limit, common to all investors. & \texteuro{}/MW/day \\
\hline
\multicolumn{3}{@{}l}{\textbf{Parameters}} \\
\hline
\(C_g\) & Marginal generation cost of generator \(g\). & \texteuro{}/MWh \\
\(\alpha,\beta\) & Intercept and slope of the affine inverse-demand (curtailment) curve; marginal willingness to pay to avoid shedding is \(\alpha+\beta\,P^{shed}_{n,t}/D^{el}_{n,t}\). & \texteuro{}/MWh \\
\(P^{max}_{g,t}\) & Maximum available dispatch of generator \(g\) in time period \(t\). & MW \\
\(Limit_l\) & Transmission capacity limit on line \(l\). & MW \\
\(PTDF_{l,n}\) & Power transfer distribution factor from node \(n\) to line \(l\). & dimensionless \\
\(D^{el}_{n,t}\) & Electricity demand at node \(n\) and time \(t\). & MW \\
\(\eta\) & One-way BESS charging and discharging efficiency, \(0 < \eta \leq 1\). & dimensionless \\
\(Limit^{node}_n\) & Accessible BESS power connection capacity at node \(n\). & MW \\
\(Ratio^{min}, Ratio^{max}\) & Minimum and maximum BESS energy-to-power ratios. & h \\
\(C^{power}_i, C^{energy}_i\) & Investor \(i\) power and energy capital cost. & \texteuro{}/MW, \texteuro{}/MWh \\
\(d_i\) & Investor \(i\) throughput degradation cost. & \texteuro{}/MWh \\
\(r_i, L_i\) & Investor \(i\) WACC and project lifetime; \(CRF_i=\dfrac{r_i(1+r_i)^{L_i}}{(1+r_i)^{L_i}-1}\). & --, a \\
\(s_{i,g}\) & Share of generator \(g\)'s spot rent owned by investor \(i\); \(0\) for a stand-alone merchant. & dimensionless \\
\end{longtable}
\endgroup

\section{Model Direction}

\noindent The extension now returns to a spot-market EPEC. The central research object is competition for scarce accessible BESS power capacity at each node, followed by strategic spot-market operation through the lower-level DC-OPF market-clearing problem. Investor heterogeneity should no longer be limited to WACC. A later formulation should distinguish investors by operational constraints, for example whether the BESS is operated as a stand-alone asset or as part of a larger portfolio.

\subsection*{Lower-Level Problem --- Spot-Market Clearing}

\paragraph{Objective}
\begin{align*}
\min \quad
&\sum_{t,g} C_g P_{g,t}
\;+\;
\sum_{n,t}
\left(
\alpha \, P^{shed}_{n,t}
+ \frac{\beta}{2 D^{el}_{n,t}} \left(P^{shed}_{n,t}\right)^2
\right)
\end{align*}
\nopagebreak[4]

\noindent The curtailment term replaces the earlier linear value-of-lost-load penalty by a strictly convex cost, i.e.\ an affine inverse-demand curve with marginal willingness to pay \(\alpha+\beta\,P^{shed}_{n,t}/D^{el}_{n,t}\). Wherever curtailment is interior this pins the nodal price to the curve, which makes the lower-level dual (and hence the price \(\lambda_{n,t}\)) unique there and bounds price spikes without a hard VOLL cap.

\paragraph{Subject to}
\begin{align*}
&\sum_{g \in G_n} P_{g,t}
  + \sum_i \left(P^{discharge}_{i,n,t} - P^{charge}_{i,n,t}\right)
  + P^{shed}_{n,t}
  - D^{el}_{n,t}
  = NetInjection_{n,t},
  &&\forall n,t
  \quad (\lambda_{n,t})
\\
\intertext{This defines nodal net injections from generation, storage dispatch, load shedding, and demand.}
&\sum_n NetInjection_{n,t}=0,
  &&\forall t
  \quad (\lambda_{sys,t})
\\
\intertext{This enforces system-wide power balance.}
&0 \le P_{g,t} \le P^{max}_{g,t},
  &&\forall g,t
  \quad (\nu^-_{g,t},\nu^+_{g,t})
\\
\intertext{This keeps conventional and renewable generation within available capacity.}
&-Limit_l
  \le
  \sum_n PTDF_{l,n} NetInjection_{n,t}
  \le
  Limit_l,
  &&\forall l,t
  \quad (\mu^-_{l,t},\mu^+_{l,t})
\\
\intertext{This represents network feasibility through PTDF-based transmission limits.}
&0 \le P^{charge}_{i,n,t} \le X^{power}_{i,n},
  &&\forall i,n,t
  \quad (\rho^+_{i,n,t})
\\
&0 \le P^{discharge}_{i,n,t} \le X^{power}_{i,n},
  &&\forall i,n,t
  \quad (\sigma^+_{i,n,t})
\\
\intertext{These constraints limit storage charging and discharging by installed power capacity.}
&SOC_{i,n,t}
  =
  SOC_{i,n,t-1}
  + \eta P^{charge}_{i,n,t}
  - \frac{1}{\eta} P^{discharge}_{i,n,t},
  &&\forall i,n,t
  \quad (\gamma_{i,n,t})
\\
\intertext{This updates the BESS state of charge from spot-market operation only.}
&0 \le SOC_{i,n,\tau} \le X^{energy}_{i,n},
  &&\forall i,n,\tau
  \quad (\delta^-_{i,n,\tau},\delta^+_{i,n,\tau})
\\
&SOC_{i,n,0}=SOC_{i,n,24},
  &&\forall i,n
\\
\intertext{These constraints keep stored energy within installed capacity and impose daily SOC consistency.}
&0 \le P^{shed}_{n,t} \le D^{el}_{n,t},
  &&\forall n,t
  \quad (\xi^-_{n,t},\xi^+_{n,t})
\end{align*}

\bigskip
\noindent\textit{Exogenous parameters fixed from the upper level or input data:}
\begin{align*}
&X^{power}_{i,n},\; X^{energy}_{i,n},\; C_g,\; P^{max}_{g,t},\;
D^{el}_{n,t},\; Limit_l,\; PTDF_{l,n},\; \eta,\; VOLL.
\end{align*}

\subsection*{Upper-Level Investment Structure}

\noindent Each investor chooses nodal BESS power and energy capacity before the lower-level spot market clears. The shared nodal access constraint is the core competition mechanism:
\begin{align*}
&\sum_i X^{power}_{i,n} \le Limit^{node}_n,
  &&\forall n.
\end{align*}

\noindent A practical storage-sizing envelope should be retained:
\begin{align*}
&Ratio^{min} X^{power}_{i,n}
\le
X^{energy}_{i,n}
\le
Ratio^{max} X^{power}_{i,n},
&&\forall i,n.
\end{align*}

\noindent Each investor maximizes deterministic spot-market profit from its BESS arbitrage and its owned generation, net of degradation and annualized CAPEX:
\begin{align*}
\max \; NPV_i
=&
\sum_{t,n}
\pi_{n,t}
\left(P^{discharge}_{i,n,t}-P^{charge}_{i,n,t}\right)
+ GenRent_i
- Cost^{degrad}_i
- CAPEX^{daily}_i,
\\[2pt]
GenRent_i
=&
\sum_{g}\; s_{i,g}
\sum_{t}
\left(\pi_{n(g),t}-C_g\right) P_{g,t},
\\[2pt]
Cost^{degrad}_i
=&
\tfrac{1}{2} d_i
\sum_{t,n}
\left(P^{charge}_{i,n,t}+P^{discharge}_{i,n,t}\right),
\\[2pt]
CAPEX^{daily}_i
=&
\frac{CRF_i}{365.25}
\sum_{n}
\left(C^{power}_i X^{power}_{i,n}+C^{energy}_i X^{energy}_{i,n}\right).
\end{align*}

\noindent The settlement price \(\pi_{n,t}\) is, by default, the nodal marginal price \(\lambda_{n,t}\) (locational marginal pricing). The implementation also supports a uniform single-zone settlement in which \(\pi_{n,t}=\lambda_{sys,t}\) for all nodes, used to study how bidding-zone pricing distorts siting relative to nodal pricing.

\noindent Investor heterogeneity enters through WACC, CAPEX, degradation, and business model. A \emph{stand-alone merchant} investor operates only BESS. A \emph{portfolio-backed} investor additionally owns a share \(s_{i,g}\in[0,1]\) of one or more of the exogenous generators already present in the lower level and collects the corresponding inframarginal spot rent \(GenRent_i\), where \(MC_g\) is the generator marginal cost, \(P_{gen,g,t}\) its cleared output, and \(n(g)\) its node; renewables have \(C_g=0\). Ownership shares are exogenous and enter only the upper-level objective, so the lower-level clearing is unchanged. Since the price \(\pi_{n,t}\) and dispatch \(P_{g,t}\) are shared with the lower level, a portfolio investor anticipates how its own BESS siting shifts the price its generation earns --- for example a solar-heavy investor time-shifting midday PV output to lift its otherwise near-zero midday nodal price.

\subsection*{Lower-Level Embedding (MPEC)}

\noindent Because the lower level is a convex quadratic program, each investor's problem is closed as a single-level MPEC by replacing the lower level with its optimality conditions, embedded through a primal--dual (strong-duality) reformulation:
\begin{itemize}
\item \emph{Primal feasibility:} all lower-level constraints above.
\item \emph{Dual feasibility:} the stationarity conditions in the dual variables listed in the notation, together with the sign restrictions on the inequality multipliers.
\item \emph{Strong duality:} the lower-level primal objective equals its dual objective, which for the convex QP is equivalent to complementary slackness and forces the embedded \((P,\lambda)\) pair onto the lower-level optimal set.
\end{itemize}
\noindent The investor's installed capacities \(X^{power}_{i,n},X^{energy}_{i,n}\) and the price \(\pi_{n,t}\) enter this MPEC simultaneously, so the investor anticipates the market-clearing response of its own investment.

\paragraph{Optimistic dual selection.} At congested or curtailed node-hours the lower-level dual is not unique, and the maximization is free to select the most favorable optimal price from that set. The embedded (``belief'') prices therefore differ from the neutral clearing prices, and the difference is reported as a diagnostic (see below). Both a generation-side and a storage-side Tikhonov regularization of the lower level were tested to remove this freedom; both either distorted the merit order/equilibrium or failed to shrink the gap at an economically negligible magnitude, so no primal regularization is used in the reported model.

\subsection*{Equilibrium Solution and Price Settlement}

\noindent The EPEC is solved by diagonalization over the investors. In each pass an investor solves its MPEC while every rival is frozen at its current capacities and represented as one aggregated, non-strategic storage unit inside the lower-level clearing. Updates use either Gauss--Jacobi (all investors respond to the same previous iterate) or Gauss--Seidel (sequential); capacity iterates are damped, and after each pass the joint capacities are projected onto the shared nodal access limit \(\sum_i X^{power}_{i,n}\le Limit^{node}_n\), preserving each investor's energy-to-power ratio.

\paragraph{Settlement and diagnostics.} Reported quantities are \emph{not} the embedded belief prices. Once the diagonalization converges, the market is cleared once more with all installed fleets fixed, and profits are settled at the resulting neutral prices \(\pi_{n,t}\) (the dual of the nodal balance for LMP settlement, or of the system balance for single-zone settlement). For each investor the model reports settled profit, the belief-minus-settled gap, and the maximum absolute difference between the embedded MPEC prices and the neutral clearing prices, so the effect of optimistic dual selection is quantified rather than hidden.

\paragraph{Robustness check against a non-strategic benchmark.} Non-unique lower-level duals raise a natural concern: if the belief price each investor embeds is not unique, is the converged capacity vector itself an artifact of which price happened to be selected, rather than a genuine equilibrium outcome? This is addressed by comparing the converged EPEC to a non-strategic benchmark that solves the identical lower-level physics --- generation dispatch, curtailment, PTDF network, storage/SOC dynamics --- and the identical shared nodal access limit and \(Ratio^{min}/Ratio^{max}\) envelope, but chooses every \(X^{power}_{i,n},X^{energy}_{i,n}\) jointly to minimize total system resource cost instead of as competing profit-maximizers. Two quantities are reported from this comparison: (i) the aggregate gap, the relative difference between total EPEC and total planner installed power/energy, \(\big|\sum_{i,n}X^{power}_{i,n}-\sum_n X^{power,\text{planner}}_n\big|/\sum_n X^{power,\text{planner}}_n\); and (ii) the siting divergence, an \(L^1\) distance between the two nodal allocation vectors after normalizing both to the same total, isolating \emph{where} capacity is placed from \emph{how much} is built. A small aggregate gap together with a non-trivial siting divergence is the expected signature when the underlying degeneracy is spatial: by standard LP duality, multiple primal optima (several nodes achieving essentially the same system cost, as evidenced by the near-zero spread in their settled hourly prices) are the counterpart of a degenerate dual, i.e.\ of the very price non-uniqueness that optimistic dual selection exploits. Under this reading, optimistic dual selection is expected to bias \emph{which} node and \emph{which investor} the capacity is attributed to, and each investor's profit level, but not the total quantity the system supports, which is instead pinned by the point where system-wide marginal storage value equals marginal storage cost.

\subsection*{Planned Extension: Nodal Access Pricing}

\noindent The diagonalization above enforces \(\sum_i X^{power}_{i,n}\le Limit^{node}_n\) by giving each investor a private, iterate-dependent headroom bound and projecting the joint sum back onto the limit whenever it is exceeded. Because that headroom depends on which investors have already updated within the current pass, the resulting split is partly an artifact of solve order rather than of relative value. The planned extension replaces the private per-investor bound with one shared node price \(\lambda^{cap}_n\), so access is allocated by willingness to pay instead of by iteration order, while leaving the aggregate limit itself unchanged.

\noindent Two changes to the upper-level problem:
\begin{align*}
&\text{(i) } X^{power}_{i,n} \text{ is no longer bounded by } Limit^{node}_n; \text{ only the } Ratio^{min}/Ratio^{max} \text{ envelope remains.}\\[2pt]
&\text{(ii) } NPV_i \;\longrightarrow\; NPV_i - \sum_n \lambda^{cap}_n X^{power}_{i,n}.
\end{align*}

\noindent The shared constraint is enforced only in aggregate, through the market-clearing condition
\begin{align*}
0 \;\le\; \lambda^{cap}_n \;\perp\; \Big(Limit^{node}_n - \sum_i X^{power}_{i,n}\Big) \;\ge\; 0,
\qquad \forall n,
\end{align*}
\noindent i.e.\ complementary slackness: a node with spare capacity clears at \(\lambda^{cap}_n=0\); a node where aggregate desired capacity would exceed \(Limit^{node}_n\) clears at the price that brings aggregate demand down to the limit exactly. \(\lambda^{cap}_n\) is found by an outer per-node price search wrapped around the existing investor diagonalization, replacing the current post-hoc projection step. As a direct by-product this traces out each investor's marginal willingness-to-pay curve \(X^{power}_{i,n}(\lambda^{cap}_n)\) at every node, which the projection-based mechanism cannot reveal.

\section*{Nine-Bus Test Network}

% Geometry-only network skeleton. Electrical data and component annotations are
% intentionally omitted.
\begin{figure}[H]
\centering
\resizebox{\linewidth}{!}{%
\begin{tikzpicture}[
  x=1cm,
  y=1cm,
  transmission/.style={
    draw=black,
    line width=0.75pt,
    line cap=round,
    line join=round
  },
  crossing/.style={
    transmission,
    preaction={draw=white,line width=3.0pt,line cap=round}
  },
  busbar/.style={
    draw=black,
    line width=2.4pt,
    line cap=butt
  },
  bus label/.style={
    font=\small,
    inner sep=1pt
  },
  line label/.style={
    font=\footnotesize,
    fill=white,
    inner sep=1.2pt
  },
  generator/.style={
    circle,
    draw=black,
    fill=white,
    line width=0.9pt,
    minimum size=0.58cm,
    inner sep=0pt,
    font=\small
  },
  load arrow/.style={
    draw=black,
    line width=0.75pt,
    -{Latex[length=4.5pt,width=3.4pt]}
  },
  component label/.style={
    font=\small,
    inner sep=1pt
  }
]

% ---------------------------------------------------------------------------
% Adjustable bus coordinates, arranged to match the reference skeleton.
% The upper buses are vertical; the middle and lower buses are horizontal.
% ---------------------------------------------------------------------------
\coordinate (n2) at (-6.00, 3.00);
\coordinate (n7) at (-3.80, 3.00);
\coordinate (n8) at ( 0.00, 3.00);
\coordinate (n9) at ( 3.80, 3.00);
\coordinate (n3) at ( 6.00, 3.00);
\coordinate (n5) at (-2.60, 0.60);
\coordinate (n6) at ( 2.60, 0.60);
\coordinate (n4) at ( 0.00,-1.20);
\coordinate (n1) at ( 0.00,-3.00);

% Generator symbol coordinates.
\coordinate (g2) at (-7.10, 3.00);
\coordinate (g3) at ( 7.10, 3.00);
\coordinate (g1) at ( 0.00,-4.15);

% ---------------------------------------------------------------------------
% Transmission connections (drawn before the busbars)
% ---------------------------------------------------------------------------

% Upper spine and the two outer links.
\draw[transmission] (n2) -- (n7); % N2--N7
\draw[transmission] (n7) -- (n8); % N7--N8
\draw[transmission] (n8) -- (n9); % N8--N9
\draw[transmission] (n9) -- (n3); % N9--N3

% Left and right descending sides of the reference skeleton.
\draw[transmission]
  (n5) -- (-2.60,2.40) -- ($(n7)+(0,-0.60)$); % N7--N5
\draw[transmission]
  ($(n5)+(0.80,0)$) -- ($(n4)+(-1.80,0)$); % N5--N4
\draw[transmission]
  ($(n4)+(1.80,0)$) -- ($(n6)+(-0.80,0)$); % N4--N6
\draw[transmission]
  (n6) -- ( 2.60,2.40) -- ($(n9)+(0,-0.60)$); % N6--N9

% Lower central link; transformer symbols are intentionally omitted.
\draw[transmission] (n4) -- (n1); % N4--N1

% Generator connections.
\draw[transmission] (n2) -- (g2);
\draw[transmission] (n3) -- (g3);
\draw[transmission] (n1) -- (g1);

% Load connections: C at Bus 8, A at Bus 5, and B at Bus 6.
\draw[load arrow]
  ($(n8)+(0,0.68)$) -- ($(n8)+(0.95,0.68)$);
\draw[load arrow]
  ($(n5)+(-0.65,0)$) -- ($(n5)+(-0.65,-0.85)$);
\draw[load arrow]
  ($(n6)+(0.65,0)$) -- ($(n6)+(0.65,-0.85)$);

% ---------------------------------------------------------------------------
% Busbars (drawn last so they remain visually dominant)
% ---------------------------------------------------------------------------

% Standard vertical busbars at the two outer buses.
\foreach \bus in {n2,n3}{
  \draw[busbar]
    ($(\bus)+(0,-0.43)$) -- ($(\bus)+(0,0.43)$);
}

% Bus 8 extends farther in the positive y direction.
\draw[busbar] ($(n8)+(0,-0.43)$) -- ($(n8)+(0,0.86)$);

% Buses 7 and 9 extend farther in the negative y direction.
\foreach \bus in {n7,n9}{
  \draw[busbar]
    ($(\bus)+(0,-0.86)$) -- ($(\bus)+(0,0.43)$);
}

% Buses 5 and 6 extend equally in both x directions.
\foreach \bus in {n5,n6}{
  \draw[busbar]
    ($(\bus)+(-1.00,0)$) -- ($(\bus)+(1.00,0)$);
}

% Bus 4 spans beneath Buses 5 and 6 so both connections can be vertical.
\draw[busbar] ($(n4)+(-2.80,0)$) -- ($(n4)+(2.80,0)$);

% Bus 1 retains the standard horizontal length.
\draw[busbar] ($(n1)+(-0.62,0)$) -- ($(n1)+(0.62,0)$);

% ---------------------------------------------------------------------------
% Bus names
% ---------------------------------------------------------------------------
\node[bus label] at ($(n2)+(0,-0.68)$) {Bus 2};
\node[bus label] at ($(n7)+(-0.55,-0.68)$) {Bus 7};
\node[bus label] at ($(n8)+(0,-0.68)$) {Bus 8};
\node[bus label] at ($(n9)+(0.48,-0.58)$) {Bus 9};
\node[bus label] at ($(n3)+(0,-0.68)$) {Bus 3};
\node[bus label] at ($(n5)+(-1.65,0)$) {Bus 5};
\node[bus label] at ($(n6)+( 1.65,0)$) {Bus 6};
\node[bus label] at ($(n4)+(-3.28,0)$) {Bus 4};
\node[bus label] at ($(n1)+(-1.05,0)$) {Bus 1};

% ---------------------------------------------------------------------------
% Line names
% ---------------------------------------------------------------------------
\node[line label] at (-4.90,3.34) {Line 2--7};
\node[line label] at (-1.90,3.34) {Line 7--8};
\node[line label] at ( 1.90,3.34) {Line 8--9};
\node[line label] at ( 4.90,3.34) {Line 3--9};
\node[line label,rotate=90] at (-2.88,1.50) {Line 5--7};
\node[line label,rotate=90] at (-2.08,-0.30) {Line 4--5};
\node[line label,rotate=90] at ( 2.08,-0.30) {Line 4--6};
\node[line label,rotate=90] at ( 2.88,1.50) {Line 6--9};
\node[line label,rotate=90] at ( 0.30,-2.10) {Line 1--4};

% Synchronous-generator symbols and names.
\node[generator] at (g2) {$\sim$};
\node[generator] at (g3) {$\sim$};
\node[generator] at (g1) {$\sim$};
\node[component label] at ($(g2)+(0,-0.55)$) {G2};
\node[component label] at ($(g3)+(0,-0.55)$) {G3};
\node[component label] at ($(g1)+(0,-0.55)$) {G1};

% Load names.
\node[component label] at ($(n8)+(1.48,0.68)$) {Load C};
\node[component label] at ($(n5)+(-0.65,-1.12)$) {Load A};
\node[component label] at ($(n6)+( 0.65,-1.12)$) {Load B};

\end{tikzpicture}%
}
\end{figure}

\end{document}
