\chapter*{Abstract}
\label{cha:abstract}
This thesis aims to determine optimal allocation strategies—specifically siting and sizing in MW and MWh—for BESS in multi-service markets, bridging the gap between technical grid constraints and economic profitability. To accurately capture the oligopolistic behavior of strategic investors, an Equilibrium Problem with Equilibrium Constraints (EPEC) framework is developed. The upper level models independent investors maximizing their individual Net Present Value (NPV), while the lower level represents an Independent System Operator (ISO) minimizing system costs via a direct current optimal power flow (DC-OPF) and an automatic Frequency Restoration Reserve (aFRR) reserve market clearing. The resulting non-linear programming (NLP) problems are solved utilizing a diagonalization algorithm. The results demonstrate that investors aggressively distribute their capacities across the entire network to exhaust all available connection limits and maximize their market shares. A major finding of this study is a severe Arbitrage-Reserve Trade-off, demonstrating that lucrative capacity premiums in the aFRR market override traditional spot market logic. The thesis concludes that evaluating technical restrictions and multi-market bidding in isolation distorts BESS investment cases. Future research should focus on integrating co-located hybrid renewable park configurations.

%\glsresetall %% all glossary entries should be used in long form (again)
%% vim:foldmethod=expr
%% vim:fde=getline(v\:lnum)=~'^%%%%\ .\\+'?'>1'\:'='
%%% Local Variables:
%%% mode: latex
%%% mode: auto-fill
%%% mode: flyspell
%%% eval: (ispell-change-dictionary "en_US")
%%% TeX-master: "main"
%%% End:
